---
title: "Lab 8 - Linear Regression in R"
author: "Camila Bobroff"
date: "11/20/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)

salmon <- read_csv("salmon_imports_new.csv")
```

**Exploratory scatterplot**
If we are going to use linear regression, we should have reason to think that a linear relationship exists between variables.

```{r}
salmon_scatter <- ggplot(salmon, aes(x = year, y = salmon_fresh))+
  geom_point()

salmon_scatter
#Overall, it looks like the relationship between year and fresh salmon imports is linear. However, the two dips in the scatterplot (around 2004 and 2014) cause some concern. It looks like it might make sense to use a linear regression in this case.
```

**Linear regression in R**
```{r}
salmon_model <- lm(salmon_fresh ~ year, data = salmon)

salmon_model
# Model equation : y = β0 + β1x + ε
# Imports(1000 lbs) = -14,982,940 + 7550(year)
```
The slope means that with every year, salmon imports are increasing by 7550 (1000 lbs).

The y-intercept means that at year 0, there were -14,982,940 (1000 lbs) of salmon being imported into the US.. this is non-sensical. Extrapolating this model for past values is not useful!

**Model diagnostics**
Use plot(model_name) to view model diagnostics in the ‘Plots’ tab (press Enter in the Console to contine to subsequent diagnostic plots). Explore the diagnostic plots to answer the following:
- Do residuals appear normally distributed?
- Any concerns about heteroscedasticity or outliers?
```{r}
par(mfrow = c(2,2)) # View all four diagnostic plots at once (and have them appear in the knitted document)
plot(salmon_model)

# The residuals appear to be normally distributed, given the qq plot. There does not seem to be an issue with heteroscedasticity, but there are a few outliers (labeled in the plots below, specifically observation 25) that might be of concern. 
```

**Explore model fit and significance**
```{r}
summary(salmon_model) # see detailed model information, including model fit information (R2 and coefficient standard errors) and statistics

# Does year significantly predict salmon imports?
# Year does significantly predict salmon imports (p <.001)

# What does the R2 value actually mean in words?
# According to our model, there is a 0.94 (R-square) proportion of variance in salmon imports that can be explained by year.

# What proportion of the variance in salmon imports is NOT explained by year?
# 1 - 0.94 = .06 of the variance in salmon imports is NOT explained by year

```

