---
title: "Lab 8 - Linear Regression in R"
author: "Camila Bobroff"
date: "11/20/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)

salmon <- read_csv("salmon_imports_new.csv")
```

**Exploratory scatterplot**
If we are going to use linear regression, we should have reason to think that a linear relationship exists between variables.

```{r}
salmon_scatter <- ggplot(salmon, aes(x = year, y = salmon_fresh))+
  geom_point()

salmon_scatter
#Overall, it looks like the relationship between year and fresh salmon imports is linear. However, the two dips in the scatterplot (around 2004 and 2014) cause some concern. It looks like it might make sense to use a linear regression in this case.
```

**Linear regression in R**
```{r}
salmon_model <- lm(salmon_fresh ~ year, data = salmon)

salmon_model
# Model equation : y = β0 + β1x + ε
# Imports(1000 lbs) = -14,982,940 + 7550(year)
```
The slope means that with every year, salmon imports are increasing by 7550 (1000 lbs).

The y-intercept means that at year 0, there were -14,982,940 (1000 lbs) of salmon being imported into the US.. this is non-sensical. Extrapolating this model for past values is not useful!

**Model diagnostics**
Use plot(model_name) to view model diagnostics in the ‘Plots’ tab (press Enter in the Console to contine to subsequent diagnostic plots). Explore the diagnostic plots to answer the following:
- Do residuals appear normally distributed?
- Any concerns about heteroscedasticity or outliers?
```{r}
par(mfrow = c(2,2)) # View all four diagnostic plots at once (and have them appear in the knitted document)
plot(salmon_model)

# The residuals appear to be normally distributed, given the qq plot. There does not seem to be an issue with heteroscedasticity, but there are a few outliers (labeled in the plots below, specifically observation 25) that might be of concern. 
```

**Explore model fit and significance**
```{r}
summary(salmon_model) # see detailed model information, including model fit information (R2 and coefficient standard errors) and statistics

# Does year significantly predict salmon imports?
# Year does significantly predict salmon imports (p <.001)

# What does the R2 value actually mean in words?
# According to our model, there is a 0.94 (R-square) proportion of variance in salmon imports that can be explained by year.

# What proportion of the variance in salmon imports is NOT explained by year?
# 1 - 0.94 = .06 of the variance in salmon imports is NOT explained by year

```

**Prepare a final graph**
Create a final graph of the linear regression model with the original data, model and 95% confidence interval for predictions.

```{r}
salmon_final_graph <- ggplot(salmon, aes(x = year, y = salmon_fresh)) + 
  geom_point() +
  geom_smooth(method = lm, se = TRUE, size = 0.5, color = "gray20") + # "Basically the CI ribbon indicates the "range" in which our predictions would be if we were to repeat the experiment (sampling) over and over. One sampling leads to a single straight line of predictions; taking into account variability of the data, the zones indicate a range of possible straight lines, if you will. By default (level at .95), we say "if we were to repeat the sampling over and over, 95% of the regression lines would be inside that grey zone."" Level can be changed by adding level = 0.99, for example, to geom_smooth.
  theme_bw() +
  scale_x_continuous(expand = c(0,0), limits = c(1990, 2017)) +
  labs(x = "Year", y = "U.S. Annual Fresh Salmon Imports\n(1,000 pounds)")

salmon_final_graph
```

**Make predictions using the model**
Let’s say we want to use our model to make predictions for U.S. fresh salmon imports in years 2022, 2024, 2026 and 2028.
```{r}
#Create a data frame containing the ‘new values’ that you want to make predictions for, feed that into the predict() function

#Make a data frame called ‘new_years’ containing the sequence of years that we want to make predictions for. Notice that the column name we create is the SAME as the variable name ‘year’ that the model uses.

new_years <- data.frame(year = c(2022, 2024, 2026, 2028))

#Then feed that data frame into the predict() function, letting it know which model to use to make predictions for the new values. Include interval = “confidence” to also report the lower and upper bounds of the 95% CI for model fit at each year.

future_predict <- predict(salmon_model, newdata = new_years, interval = "confidence")
future_predict

#Bind the prediction outcomes (those are the values in column ‘fit’) with the ‘new_years’ data to actually create a useful table of predicted imports and upper and lower CI:

predictions <- data.frame(new_years, future_predict)
predictions

```

